\section{RQ1: Accuracy Evaluation of Unused Dependency Detection Models}
\label{sec:rq1}

\subsection{Motivation}

Unused dependency detection is the foundation of our skip analysis.
If a model incorrectly classifies a \emph{used} dependency as unused, a
build that should run would be skipped, potentially concealing a breakage.
Conversely, if the model misses genuinely unused dependencies, the potential
CI savings are underestimated.
It is therefore essential to quantify the precision and recall of each of
the five detection models before interpreting downstream results.

\subsection{Approach}

\subsubsection{Project Selection}

For RQ1, we use the projects that passed PS8 (Section~\ref{sec:ps8}),
i.e., projects with a passing test suite and a line coverage of at least
70\% as measured by nyc.
Because these projects have a reliable and well-exercised test suite, we
can use test outcomes as a ground truth oracle for dependency usage:
\emph{a dependency is truly unused if and only if removing it does not
cause any test to fail}.

\subsubsection{Reference Set Construction}
\label{sec:ref-set}

To determine which dependencies are truly unused in each project,
we attempt to remove each dependency listed in \texttt{package.json}
independently.
For each candidate dependency $d$:

\begin{enumerate}
  \item Run \texttt{npm uninstall $d$} in the cloned repository.
  \item Execute the nyc-instrumented test command established in PS8.
  \item If all tests pass (exit code 0), record $d$ as \emph{truly unused}.
  \item Restore the original \texttt{package.json} and
        \texttt{node\_modules} before proceeding to the next candidate.
\end{enumerate}

\noindent The set of all truly unused dependencies identified across all
PS8 projects is used as the reference for computing recall.

\subsubsection{Unused Dependency Detection}

We applied five detection models to each PS8-passing project.
For static analysis, we used depcheck, which runs \texttt{depcheck . --json},
and knip, which runs \texttt{npx knip --reporter json}.
For LLM-based detection, we used three models served locally via
the Ollama HTTP API: llama3.1:8b, qwen3.5:4b, and
deepseek-coder:6.7b-instruct.
These three models were selected because they are lightweight,
open-source, and widely recognised in the community,
making them suitable for local inference on an HPC cluster without
relying on external API services.
For each LLM, we extracted all \texttt{import}/\texttt{require}
statements from the source files and passed them, together with the
\texttt{package.json} dependency list and relevant configuration files,
as a prompt to the model.

\subsubsection{Evaluation}

For each model, we take every dependency it predicted as unused and
remove it from the project using \texttt{npm uninstall}.
We then re-execute the nyc-based test command established in PS8 and
record whether the tests pass or fail.
If the tests still pass after removal, the prediction is correct
(the dependency was truly unused); if the tests fail, the prediction
was wrong (the dependency was actually used).
This gives us the number of correct predictions (true positives, TP)
and incorrect predictions (false positives, FP) for each model.

Dependencies that were confirmed as truly unused by the removal procedure
described above, but were not predicted by the model, are counted as
false negatives (FN).
Precision and recall are then defined as:

\begin{equation}
  \text{Precision} = \frac{|\mathrm{TP}|}{|\mathrm{TP}| + |\mathrm{FP}|},
  \qquad
  \text{Recall} = \frac{|\mathrm{TP}|}{|\mathrm{TP}| + |\mathrm{FN}|}
\end{equation}

\subsection{Results}
% TODO: Fill in after experiments complete.

\subsection{Summary}
% \begin{mdframed}
% \textbf{RQ1 Summary:} ...
% \end{mdframed}
