\section{RQ2: Quantitative Evaluation of CI Skippability}
\label{sec:rq2}

\subsection{Motivation}

Even when a CI build passes, it may have been unnecessary if the only
change was a version bump of a dependency that is not actually used in
the project.
RQ2 aims to quantify how often such unnecessary builds occur and how
much CI execution time they consume.

\subsection{Approach}

\paragraph{DC1: Dependency-Update Commit Extraction.}
From the git history of projects that passed PS6, we extracted commits
that modified \emph{only} \texttt{package.json} and/or
\texttt{package-lock.json} and contained exactly one dependency version
update.
Commits that also modified source files, test files, README files, or
any other artifact were excluded.
This strict criterion ensures that each extracted commit can be
attributed to a single dependency version bump.
We note that this exclusion criterion may omit commits that update an
unused dependency while simultaneously making trivial changes to other
files (e.g., adding a comment or updating the README).
Consequently, the waste reported in this study represents a conservative
lower bound of the actual CI waste induced by unused dependency updates.

\paragraph{DC2: CI Data Collection.}
For each extracted commit and its parent commit, we queried the GitHub
Check Runs API to obtain the CI execution status
(\texttt{success} / \texttt{failure}) and the total runtime in seconds.
CI data could be retrieved for XX\% of the extracted commits; the
remaining commits were excluded from further analysis because their
CI outcomes could not be determined.

\paragraph{DC3: Unused Dependency Detection.}
For commits with CI data available for both the commit and its parent,
we cloned each repository and checked out the target commit.
We then ran the five detection models on the checked-out source tree.

\paragraph{Skip Analysis.}
A build is classified as skippable if and only if:

\begin{equation}
  \textit{skippable} \;\Leftrightarrow\;
  \bigl(\textit{updated\_dep} \in \textit{unused\_dep}
        \cup \textit{unused\_dev\_dep}\bigr)
  \;\wedge\;
  \bigl(\textit{parent\_build\_status} = \texttt{success}\bigr)
\end{equation}

\noindent The parent-build condition ensures that we only consider
skipping builds whose pre-update state was already passing.
For each model, we calculated the proportion of skippable builds and
the total CI execution time that could have been saved.

\subsection{Results}
% TODO: Fill in after experiments complete.

\subsection{Summary}
% \begin{mdframed}
% \textbf{RQ2 Summary:} ...
% \end{mdframed}
